\section*{Homework 10 Due to April 11 8:00 AM}
\question{1}{}
Assume we have Yang-Mills theory with the gauge sector
\begin{align}
    G=SU(2),
\end{align}
and the Higgs sector consisting of one real scalar field in the \textit{adjoint} representation of $SU(2)$.

The adjoint field develops a vacuum expectation value $\langle\phi^a\phi^a\rangle=v^2$. Calculate the first coefficient in the gauge coupling $\beta$ function for $\mu\gg M_W, M_H$. Compare the result to the first coefficient in the $\beta$ function of pure Yang-Mills theory with the same gauge group as above. Can you give a physical interpretation for the difference?

Assume that $M_W \ll M_H$. Find the first coefficients in the $\beta$ function in
the window $M_W \ll \mu \ll M_H$.

\answer{}
We quote the equation (16.133) in Peskin:
\begin{align}
\beta(g)=-\frac{g^3}{(4\pi)^2}\left(\frac{11}{3}C_2(G)-\frac{4}{3}n_f C(r)\right).
\end{align}
However, we can generalize this equation by including the Higgs sector. For a theory with scalars in the representation $r_s$, the first coefficient $b_0$ of the $\beta$ function (where $\beta(g) = -\frac{b_0}{16\pi^2}g^3$) is given by:
\begin{align}
b_0 =& \frac{11}{3}C_2(G) - \sum_{s} \frac{1}{3} C(r_s) \cdot (\text{const}) - \sum_{f} \frac{4}{3} S_2(r_f)\\
=&\frac{11}{3}C_2(G) - \sum_{s} \frac{1}{3} C(r_s) \cdot (\text{const})
\end{align}
where $C(r_s)$ is the index of the representation $r_s$ (often denoted as $T(r_s)$), while the spin-dependent factors are already included in the prefactors of the formula. For a real scalar field in the adjoint representation of $G=SU(2)$, we have:
\begin{itemize}
\item $C_2(G) = 2$.
\item $C(adj) = C_2(G) = 2$.
\item For a real scalar, the contribution is half of that of a complex scalar. Thus, the coefficient is $1/6$ instead of $1/3$ in the standard complex scalar formula. More specificly, the const in the equation now is $\frac{1}{2}$.
\end{itemize}

\begin{itemize}
    \item [(a)] $\mu\gg M_W,M_H$: 
\end{itemize}
In this regime, we can treat all fields in the theory as effectively massless. Both the gauge bosons and the adjoint scalar field contribute to the vacuum polarization. Using the values identified above, the first coefficient $b_0$ is: 
\begin{align}
b_0^{\text{high}} = \frac{11}{3}(2) - \frac{1}{6}(2) = \frac{22}{3} - \frac{1}{3} = \frac{21}{3} = 7.
\end{align}

\begin{itemize}
    \item [(b)]  $M_W \ll \mu \ll M_H$:
\end{itemize}
Since we assume $M_W \ll \mu \ll M_H$, the scalar field (with mass $M_H$) has "frozen out" or decoupled from the renormalization group flow. However, the gauge bosons remain active since $\mu$ is still much larger than their mass scale $M_W$. Therefore, the theory effectively behaves like a pure $SU(2)$ Yang-Mills theory in this regime. The first coefficient is:
\begin{align}
    b_0^{\text{mid}} = \frac{11}{3}C_2(G) = \frac{11}{3}(2) = \frac{22}{3}.
\end{align}
In this window, the coupling $g(\mu)$ runs faster (the beta function is more negative) than in the high-energy regime because the screening contribution from the adjoint scalar is no longer present.


\textbf{Physical Interpretation of the difference:} Comparing $b_0^{\text{high}} = 7$ to $b_0^{\text{mid}} = 22/3$, we see that the presence of the adjoint scalar field decreases the value of $b_0$. In Yang-Mills theory, the gauge boson self-interactions lead to anti-screening (asymptotic freedom), which corresponds to the $11/3$ term. On the other hand, the scalar field contributes to screening (the QED-like behavior). Therefore, adding an adjoint scalar field increases the screening effect in the UV, which makes the theory "less" asymptotically free compared to the pure gauge theory.