\section*{Homework 7 Due to March 19 8:00 AM}
\question{1}{}
Compare the composition of the Dirac spinor and all five gamma matrices, including 
\begin{align}
    \gamma^5 = \pm i \prod_{\mu=0,1,2,3} \gamma^\mu,
\end{align}
in Peskin and Schroeder, Landau-Lifshitz’s Quantum Electrodynamics, and with the one introduced in my course. Find the unitary transition matrix $U$ which implements the transition from the spinorial to the standard (Dirac) representation.
\answer{}
In Peskin and Schroeder, the gamma matrices are defined as
\begin{align}
    \gamma^0 = \begin{pmatrix}
        0 & I \\
        I & 0
    \end{pmatrix}, \quad
    \gamma^i = \begin{pmatrix}
        0 & \sigma^i \\
        -\sigma^i & 0
    \end{pmatrix}, \quad
    \gamma^5 = i\gamma^0\gamma^1\gamma^2\gamma^3 = \begin{pmatrix}
        -I & 0 \\
        0 & I
    \end{pmatrix},
\end{align}
where $\sigma^i$ are the Pauli matrices. In Landau-Lifshitz’s Quantum Electrodynamics, the gamma matrices are defined as
\begin{align}
    \gamma^0 = \begin{pmatrix}
        I & 0 \\
        0 & -I
    \end{pmatrix}, \quad
    \gamma^i = \begin{pmatrix}
        0 & \sigma^i \\
        -\sigma^i & 0
    \end{pmatrix}, \quad
    \gamma^5 = -i\gamma^0\gamma^1\gamma^2\gamma^3 = \begin{pmatrix}
        0 & -I \\
        -I & 0
    \end{pmatrix}.
\end{align}
In the course, the gamma matrices are defined as
\begin{align}
    \gamma^0 = \begin{pmatrix}
        0 & I \\
        I & 0
    \end{pmatrix}, \quad
    \gamma^i = \begin{pmatrix}
        0 & \sigma^i \\
        -\sigma^i & 0
    \end{pmatrix}, \quad
    \gamma^5 = i\gamma^0\gamma^1\gamma^2\gamma^3 = \begin{pmatrix}
        -I & 0 \\
        0 & I
    \end{pmatrix}.
\end{align}
The unitary transition matrix $U$ that implements the transition from the spinorial to the standard (Dirac) representation can be found by comparing the definitions of the gamma matrices in the two representations. We can write the gamma matrices in the spinorial representation as
\begin{align}
    \gamma^0_{\text{spinorial}} = \begin{pmatrix}
        0 & I \\
        I & 0
    \end{pmatrix}, \quad
    \gamma^i_{\text{spinorial}} = \begin{pmatrix}
        0 & \sigma^i \\
        -\sigma^i & 0
    \end{pmatrix}, \quad
    \gamma^5_{\text{spinorial}} = i\gamma^0_{\text{spinorial}}\gamma^1_{\text{spinorial}}\gamma^2_{\text{spinorial}}\gamma^3_{\text{spinorial}} = \begin{pmatrix}
        -I & 0 \\
        0 & I
    \end{pmatrix}.
\end{align}
The gamma matrices in the standard (Dirac) representation are
\begin{align}
    \gamma^0_{\text{Dirac}} = \begin{pmatrix}
        I & 0 \\
        0 & -I
    \end{pmatrix}, \quad
    \gamma^i_{\text{Dirac}} = \begin{pmatrix}
        0 & \sigma^i \\
        -\sigma^i & 0
    \end{pmatrix}, \quad
    \gamma^5_{\text{Dirac}} = i\gamma^0_{\text{Dirac}}\gamma^1_{\text{Dirac}}\gamma^2_{\text{Dirac}}\gamma^3_{\text{Dirac}} = \begin{pmatrix}
        0 & I \\
        I & 0
    \end{pmatrix}.
\end{align}
To find the unitary transition matrix $U$, we can use the fact that the gamma matrices in the two representations are related by a similarity transformation:
\begin{align}
    \gamma^\mu_{\text{Dirac}} = U \gamma^\mu_{\text{spinorial}} U^\dagger.
\end{align}
By comparing the definitions of the gamma matrices in the two representations, we can find that the unitary transition matrix $U$ is given by
\begin{align}
    U = \frac{1}{\sqrt{2}} \begin{pmatrix}
        I & I \\
        -I & I
    \end{pmatrix}.
\end{align}
This matrix satisfies the relation
\begin{align}
    U \gamma^0_{\text{spinorial}} U^\dagger = \gamma^0_{\text{Dirac}}, \quad
    U \gamma^i_{\text{spinorial}} U^\dagger = \gamma^i_{\text{Dirac}}, \quad
    U \gamma^5_{\text{spinorial}} U^\dagger = \gamma^5_{\text{Dirac}}.
\end{align}
\qed


\clearpage
\question{2}{}
Compare the definitions of left-handed and right-handed. As a reference point use the Dirac equation of motion
\begin{align}
    i \gamma^\mu \partial_\mu \psi = m \psi
\end{align}
and the fact that neutrino is left-handed. Check that in all definitions of the
Dirac matrices the Clifford algebra is satisfied.
\answer{}
We can define the Dirac spinor $\psi$ as a four-component with a left-handed Weyl spinor $\psi_L$ and a right-handed Weyl spinor $\psi_R$, which have 2 components each:
\begin{align}
    \psi = \begin{pmatrix}
        \psi_L \\
        \psi_R
    \end{pmatrix}.
\end{align}
The left-handed and right-handed components of the Dirac spinor can be defined using the projection operators:
\begin{align}
    P_L = \frac{1}{2}(1 - \gamma^5), \quad P_R = \frac{1}{2}(1 + \gamma^5).
\end{align}
The left-handed and right-handed components of the Dirac spinor can be obtained by applying the projection operators to the Dirac spinor:
\begin{align}
    \psi_L = P_L \psi, \quad \psi_R = P_R \psi.
\end{align}
We can write the Dirac equation of motion in terms of the left-handed and right-handed components:
\begin{align}
    i \begin{pmatrix}
        0 & \sigma^\mu \\
        \bar{\sigma}^\mu & 0
    \end{pmatrix}
    \begin{pmatrix}
        \partial_\mu \psi_L \\
        \partial_\mu \psi_R
    \end{pmatrix} = m \begin{pmatrix}
        \psi_L \\
        \psi_R
    \end{pmatrix},
\end{align}
where $\sigma^\mu = (I, \sigma^i)$ and $\bar{\sigma}^\mu = (I, -\sigma^i)$. This equation can be separated into two equations:
\begin{align}
    i \sigma^\mu \partial_\mu \psi_R = m \psi_L, \quad
    i \bar{\sigma}^\mu \partial_\mu \psi_L = m \psi_R.
\end{align}
In the massless limit ($m=0$), the equations decouple and we have
\begin{align}
    i \sigma^\mu \partial_\mu \psi_R = 0, \quad
    i \bar{\sigma}^\mu \partial_\mu \psi_L = 0.
\end{align}
This shows that the left-handed and right-handed components of the Dirac spinor are independent in the massless limit. Hence, the massless fermion can be left-handed or right-handed. Hence, the neutrino, which is observed to be left-handed, can be described by a left-handed Weyl spinor $\psi_L$ in the massless limit.

To check that the Clifford algebra is satisfied in all definitions of the Dirac matrices, we need to verify that the gamma matrices satisfy the anticommutation relation:
\begin{align}
    \{\gamma^\mu, \gamma^\nu\} = 2g^{\mu\nu}I,
\end{align}
where $g^{\mu\nu}$ is the Minkowski metric. Here we use the Weyl representation of the gamma matrices, which is given by
\begin{align}
    \gamma^0 = \begin{pmatrix}
        0 & I \\
        I & 0
    \end{pmatrix}, \quad
    \gamma^i = \begin{pmatrix}
        0 & \sigma^i \\
        -\sigma^i & 0
    \end{pmatrix}, \quad
    \gamma^5 = i\gamma^0\gamma^1\gamma^2\gamma^3 = \begin{pmatrix}
        -I & 0 \\
        0 & I
    \end{pmatrix}.
\end{align}
Then we can compute the anticommutators:
\begin{align}
    \{\gamma^0, \gamma^0\} &= 2I, \\
    \{\gamma^i, \gamma^j\} &= -2\delta^{ij}I, \\
    \{\gamma^0, \gamma^i\} &= 0, \\
    \{\gamma^5, \gamma^\mu\} &= 0.
\end{align}
Here we apply the properties of the Pauli matrices, which satisfy $\{\sigma^i, \sigma^j\} = 2\delta^{ij}I$. Since we can apply the unitary transition matrix $U$ to transform the gamma matrices to the standard (Dirac) representation, we can conclude that the Clifford algebra is satisfied in all definitions of the Dirac matrices.
\qed