\section*{HW2 Due to October 7 11:59 PM}

\question{1}{Problem 5.1}\\
Work out the LSZ reduction formula for the complex scalar field that was introduced in problem~$3.5$. Note that we must specify the type ($a$ or $b$) of each incoming and outgoing particle.
\answer{}
\clearpage
\question{1}{Problem 6.1}
\begin{itemize}
    \item [(a)] Find an explicit formula for $\mathcal{D}q$ in eq.~(6.9). Your formula should be of the form $\mathcal{D}q = C \prod_{j=1}^N dq_j$, where $C$ is a constant that you should compute.
    \item [(b)] For the case of a free particle, $V(Q)=0$, evaluate the path integral of eq.~(7.9) explicitly. Hint: integrate over $q_1$, then $q_2$, etc, and look for a pattern. Express you final answer in terms of $q',t',q'',t''$ and $m$. Restore $\hbar$ by dimensional analysis.
    \item [(c)] Compute the $\langle q'',t''|q',t' \rangle =\langle q'' |e^{-iH(t''-t')}|q'\rangle$ by inserting a complete set of momentum eigenstates, and performing the integral over the momentum. Compare your result in part (b).
\end{itemize}
\answer{}
\clearpage
\question{1}{Problem 7.3}
\begin{itemize}
    \item [(a)]Use the Heisenberg equations of motion, $\dot{A}=i[H,A]$, to find explicit expressions for $\dot{Q}$ and $\dot{P}$. Solve these to get the Heisenberg-picture operators $Q(t)$ and $P(t)$ in terms of the Schr\"odinger-picture operators $Q$ and $P$.
    \item [(b)] Write the Schr\"odinger-picture operators $Q$ and $P$ in terms of the creation and annihilation operators $a$ and $a^\dagger$, where $H=\hbar\omega (a^\dagger a +\frac{1}{2})$. Then, using your result from part (a), write the Heisenberg-picture operator $Q(t)$ and $P(t)$ in terms of $a$ and $a^\dagger$.
    \item [(c)] Using your result from part (b), and $a|0\rangle=\langle0|a^\dagger=0$,verify eqs.~(7.16) and (7.17).
\end{itemize}


\answer{}

\clearpage
\question{1}{Problem 7.4}\\
Consider a harmonic oscillator in its ground state at $t=-\infty$. It is then subjected to an external force $f(t)$. Compute the probability $|\langle0|0\rangle_f|^2$ that the oscillator is still in its ground state at $t=+\infty$. Write your answer as a manifestly real expression, and in terms of the Fourier transform $\tilde{f}(E)=\int^{+\infty}_{-\infty}e^{iEt}f(t)$. Your answer should not involve any other unevaluated integrals.
\answer{}